Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-07-29 doc/manuscript/response_to_reviewers_original_submission.qmd

-   The authors have 7 state variables which they define as 'density' of individuals in different life history states.
    Is this really density (individuals per unit area) or just the number of individuals.

    -   [Good point. We are modeling abundance per population so we have replaced the word density with abundance to avoid confusion over the units of area.]{style="color: blue"}

-   Additionally, are the authors modelling both males and females, or just females.
    Based on the model structure it looks like they are modelling females, as there is no adjustment for the proportion of females in the population.
    This should be clarified.
    If I am correct, then shouldn't fecundity be expressed as the number of female eggs?

    -   [Thank you for the careful reading of our model. We are modeling the entire population, not just females, and are assuming a 50:50 sex ratio. This was not explicit in our equations and we have thus updated our viability model equation to explicitly include a parameter for sex ratio ($\rho$). We now describe this parameter in the Table in the supplementary material. While addressing this comment, we noticed that we were not consistent with modeling females vs. total population between the deterministic and stochastic models. We have updated the deterministic model to be exactly analogous to the stochastic model (i.e., both are now modeling total population). Our conclusions regarding viability are unchanged.]{style="color: blue"}

@@

-   I am also a little confused at what they consider 'recruitment' in this model.
    I understand the survival probability, indicated by sigma notation (σ), but none of the parameters listed in Supplementary Table 1 are termed recruitment.
    However, in both the text the term is used in multiple places, e.g., lines 106, 102, 458 and figure legends (Fig 5), but I'm sure it means different thing in these places.
    For this reason, I think the authors should clarify their use of the term, especially in the viability modelling.
    To me, recruitment is conditional on survival so I would tend to equate it with (1- 𝑝𝐽1) or the probability that given a juvenile survives it 'recruits' or transitions into an adult frog.
    I don't think that is how the adults are using it in this model.
    I think they could refrain from using it at all in this section.

    -   [Good question. This was a point of some discussion within our group. For a juvenile to recruit to an adult, it has to survive, and thus $\sigma_{J_1}$ encompasses survival and recruitment. This is why we used the language of "survival and recruitment" to describe this parameter in the text. However, we see how this could be confusing. As we do not lose much biological interpretation by just referring to $\sigma_{J_1}$ as "survival" we have made this change throughout the text as suggested by the reviewer.]{style="color: blue"}

@@

4.  Given the page limitations employed by the journal, the authors must make decisions about what information to supply.
    Personally, I would have liked to see a little more information on the variation in recruitment (number of new adults a population) and population growth (change in abundance over time) among recipient sites.
    These are estimated (line 106; Supplementary Material) but not reported from their Bayesian Jolly-Seber CMR model.
    Instead, the authors chose to use limited demographic estimates from reintroduced populations (adult survival only) in a population viability model to explore variation in population viability over 50 years.
    The only thing that varies across their 12 reintroduced populations in this exercise is adult survival; the other parameters are the same across populations (Supplementary Table 1).
    In essence, this exercise shows how population viability varies with adult survival.
    It's really important to choose reintroduction sites where adult annual survival exceeds 0.50, preferably 0.60, and the authors make a case that survival 1-year post-translocation is a good proxy for habitat quality; however, part of that case is assuming that all other demographic parameters are the same among populations, which is probably not true.
    The elasticity and sensitivity analyses associated with the viability analyses do suggest that adult survival is certainly the most important parameter (almost always true in vertebrate systems).
    Still, the decision of what to present given limited space is up to the authors and just because it does not match what this one reviewer would have preferred does not detract from the quality and quantity of work demonstrated in this paper.
    This is a really good paper and definitely worthy of publication in Nature Communications.

    -   [We appreciate the reviewer's thoughtful point. It is worth noting that the viability analysis also explores variation in $\sigma_{J_1}$ --- a parameter that our sensitivity analyses indicate is likely the second most important parameter for viability following adult survival. So we are actually exploring viability across environmental conditions that affect the two most important parameters for population growth. While we agree that other vital rates likely vary to some degree across population, we have good empirical and modeling evidence that the two parameters we explore are the most important for long-term viability. As no changes were requested we, we did not make any.]{style="color: blue"}

@@

5.  Lines 221-233.
    This is very true -- congratulations!!

    [Thanks]{style="color: blue"}

6.  Line 252.
    I suggest revising this sentence; frog survival at natal (donor) sites was never reported.
    If you are referring to the difference between translocated and naturally recruited adults at the 3 recipient locations (Supplementary Fig. 3) I would clarify that.

    [ROLAND: I am not sure what exactly this is referring to as I don't have the line numbers for the submission.]{style="color: blue"}

@@

### Reviewer 3

General comments: This is a long-term dataset, in many ways exceptionally long for in depth amphibian studies, and a valuable manuscript on a priority topic for biodiversity conservation.
Bd research has been criticised for missing direct conservation applications in the field, which makes this study particularly novel and insightful, despite focusing on a single species.
The results are very encouraging and give much cause of cautious optimism for what has often appeared to be an almost intractable problem for amphibian conservation but while I understand the specificity of your results it would be good to have at least a paragraph that looks at the potential implications for other species and important steps forward that could be concluded from your own work.
In my opinion the text is clear and very well-written and I do not have much to add in relation to that.
There is some repetition between introduction and discussion so please be aware of that.
As a general comment, you mention that the three donor populations are large, with several hundred adults and that you have documented the genetic structure but I suggest it would be useful to briefly discuss the potential implications of the reintroductions from a genetic perspective and any potential monitoring follow-up of that aspect and mitigation strategy if required.
There is a reasonable potential for creating a genetic bottleneck as a results of these reintroductions in the absence of some kind of genetic monitoring.
Also, it would be useful to briefly explain why are reintroductions important (and cost-effective) instead of simply waiting for these recovering populations to naturally expand over time.
The methods look sound but I have limited experience in this type of multilevel Bayesian modelling.

Detailed comments Title- Just a suggestion but for me the title would benefit from mentioning the species.
Your work obviously has wider implications but remains currently very species-specific.
Line 155- for me this aspect of the apparent high variability linked to donor population is important and I suggest should be highlighted both in the results and especially in the discussion as a separate point.
Would the reintroductions be feasible in terms of achieving long term population viability with the lowest survival based on donor population?

-   [Good question. Our population viability analyses show that long-term viability is significantly more likely when 1-year adult survival probability is \>0.5. We hesitate to try to separate the effect of recipient site and donor population on 1-year adult survival as these factors were not crossed in our study (an inevitable logistical constraint when working on real landscapes and taking into account host landscape genetics \[ROLAND, update as needed\]). That being said, there is strong support from the model that the lowest 1-year survival from donor population 72996 would not be sufficient for long-term viability (\< 0.5) while the lowest 1-year adult survival from 70459 and 70567 would be sufficient for long-term viability (\> 0.5). While it is tempting to conclude that donor populations 70459 and 70567 should be the focus of subsequent translocations, our design cannot disentangle site and donor effects.]{style="color: blue"}

Line 266- I suggest more caution in this statement and perhaps combining with the following sentence, which seems a reasonable explanation but remains based on a relatively small number of sites and reintroduction events.
I would strongly expect specific weather patterns to also play an important role in any given year, even if these might be rarer events that you have not yet captured.
Line 288- for me the value in this call for additional research warrants some brief but specific details about what that would entail.
Otherwise, researchers naturally always call for additional research.
Lines 298-301- I think this is important and relevant in the wider amphibian conservation context, where juveniles at this crucial life stage are often the specifically missing dataset because they are harder to monitor.
This is in many ways a failure of the sector and the opposite compared to say birds or mammals, yet has become inbuilt in much amphibian research and is worth highlighting as it has important consequences for progressing practical conservation.
Lines 310-315 I suggest you could be more forward looking here and suggest some critical steps forward for interpreting the success of your results for advancing the conservation prospects for other similarly threatened amphibian species.
What are the crucial considerations that should inform their strategies and how would you go about it?
It might also be useful to also briefly mention the relevance of your results for reintroductions from the perspective of captive breeding and amphibian arks given that the current approaches would not facilitate any kind of resistance building (if at all possible).
Line 317- why would it though?
Now that you have developed a solid understanding of the process, could you not strategically reintroduce populations to facilitate ongoing natural colonisation once there are sufficient numbers to create some wider population resilience?
Line 337- this is an important methodological point and which requires a bit more substance in terms of how was absence determined, over what period, etc

### 2024-07-29 doc/manuscript/supplementary_information.qmd

| **Parameter**                                                             | **Value**            | **Source**                                                                        |
|:-------------------------|:------------------|:---------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year-1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_2}$, Yearly survival probability of year-2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_3}$, Yearly survival probability of year-3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{J_1}$, Yearly survival probability of year-1 juveniles.          | Varies yearly        | Varies. Bounds estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year-2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from CMR studies                                                        |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from CMR studies                                                        |
| $p_{L_1}$, Probability of a year-1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge                |
| $p_{L_2}$, Probability of a year-2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_{J_1}$, Probability of a year-1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_F$, Probability of an adult female reproducing in a year               | 0.5                  | Could be as high at 1, based on field observations                                |
| $\rho$, Proportion of adults that are female                              | 0.5                  | SOURCE                                                                            |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                                |

>>>>

| **Parameter**                                                             | **Value**            | **Source**                                                                        |
|:-------------------------|:------------------|:---------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year-1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_2}$, Yearly survival probability of year-2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_3}$, Yearly survival probability of year-3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{J_1}$, Yearly survival probability of year-1 juveniles.          | Varies yearly        | Varies. Bounds estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year-2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from CMR studies                                                        |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from CMR studies                                                        |
| $p_{L_1}$, Probability of a year-1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge                |
| $p_{L_2}$, Probability of a year-2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_{J_1}$, Probability of a year-1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_F$, Probability of an adult female reproducing in a year               | 0.5                  | Could be as high at 1, based on field observations                                |
| $\rho$, Proportion of adults that are female                              | 0.5                  | Estimated from field data, observations, natural history knowledge                |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                                |

### 2024-07-29 doc/manuscript/translocation.qmd

Given observed yearly adult survival probabilities of translocated frogs (from site-specific mrmr CMR models; provided in legend of [@fig-viability]B) and a yearly survival probability of the year-1 juvenile class ($\sigma_{J_1}$) greater than 0.09, at least six of twelve translocated populations should experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd ([@fig-viability]A; median predicted $\lambda$ ranges from 1.19-1.40 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
As year-1 juvenile survival probability increased above 0.2, the deterministic long-run growth rate of eight of twelve population was greater than 1 ([@fig-viability]A).

Even when incorporating (i) demographic stochasticity and (ii) environmental stochasticity in year-1 juvenile survival and recruitment (the transition that we expect to be the most subject to environmental variability in the presence of Bd), populations with high adult survival are likely to persist over a 50 year time horizon.
Our model predicted that, following a single introduction of 40 adult individuals into a population, the six populations with the highest adult survival probabilities ($\sigma_{A_R} > 0.5$) had 50-year extinction probabilities of less than 0.5 when the average year-1 juvenile survival was greater than 0.10 ([@fig-viability]B).
This indicates strong potential for long-term persistence in the presence of Bd and environmental variability in survival and recruitment.
In contrast, for the six populations where yearly adult survival probability $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be \> 50% regardless of the value of mean year-1 juvenile survival between 0 and 0.25.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history ([@fig-viability]C; population 70550, surveyed for 16 years).

@@

The recent emergence of Bd worldwide has contributed to the decline of hundreds of amphibian species, some of which are now extinct in the wild [@scheele2019].
This extraordinary impact on global amphibian biodiversity is compounded by the lack of any effective and broadly applicable strategies to reverse these impacts [@garner2016; @knapp2022].
Importantly, in addition to the natural recovery documented for *R. sierrae* [@knapp2016], other amphibian species are also showing evidence of post-epizootic recovery in the presence of Bd [@scheele2017; @voyles2018] and suggest the possibility of also using animals from these recovering populations to reestablish extirpated populations.
As with *R. sierrae*, the feasibility and long-term success of such efforts will depend on the availability of robust donor populations containing individuals that have sufficiently high resistance to allow frog survival and population growth in the presence of Bd.
Despite the promising example provided by our study of successful reestablishment of *R. sierrae* despite ongoing Bd infection, the challenge of recovering hundreds of Bd-impacted amphibian species globally is a daunting prospect.
An adaptive management approach such as the one we utilized, in which the outcomes of recovery actions are rigorously assessed and this information is used to reduce uncertainties and inform future actions, will be essential in maximizing the effectiveness of this global recovery effort \[@canessa2019\]. In addition, developing an improved understanding of the factors facilitating or limiting the development of frog resistance against Bd infection and the effectiveness of selective breeding programs and other genetic interventions to enhance frog resistance \[@kosch2022\] may be particularly important.
Although we now have a proven strategy to reestablish extirpated *R. sierrae* populations, recovery across their large historical range will require substantial resources over many decades.
The results of this study provide a hopeful starting point for that endeavor and other future efforts worldwide.

@@

To determine the implications of observed 1-year adult survival on the long-term viability of populations established via translocation, we developed a population model for *R. sierrae*.
Our central question was: How does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations?
We developed a model that tracked seven state variables of a frog population: density of translocated adults ($A_T$), density of adults naturally recruited into the population ($A_R$), density of first-year tadpoles ($L_1$), density of second-year tadpoles ($L_2$), density of third-year tadpoles ($L_3$), density of first-year juveniles ($J_1$), and density of second-year juveniles ($J_2$).
We divided adults into two classes $A_T$ and $A_R$ because there is evidence that the survival probability of translocated adults and naturally recruited adults differs (Supplementary Fig. 3).

@@

$$
\begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\ 
A_T
\end{bmatrix}(t + 1) = 
\begin{bmatrix}
  0 & 0 & 0 & 0 & 0 & \sigma_{A_R} p_F F & \sigma_{A_T} p_F F \\
  \sigma_{L1} p_{L1} & 0 & 0 & 0 & 0 & 0 & 0 \\
  0 & \sigma_{L2} p_{L2} & 0 & 0 & 0 & 0 & 0\\
  \sigma_{L1} (1 - p_{L1}) & \sigma_{L2} (1 - p_{L2}) & \sigma_{L3} & 0 & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} p_{J1} & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} (1 - p_{J1}) & \sigma_{J2}  & \sigma_{A_R} & 0 \\
  0 & 0 & 0 & 0 & 0 & 0 & \sigma_{A_T} 
\end{bmatrix} \begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\
A_T \\
\end{bmatrix}(t) 
$$

The parameters in this model are yearly survival probability $\sigma_{\cdot}$ (the subscript "$\cdot$" indicates a particular state variable), probability that a female frog reproduces in a given year $p_F$, number of eggs produced by a female frog in a year that successfully hatch $F$, probability of a first-year tadpole remaining as a tadpole $p_{L1}$, probability of a second-year tadpole remaining as a tadpole $p_{L2}$, and probability of a first-year juvenile remaining as a juvenile $p_{J1}$.
First-year juvenile survival and recruitment $\sigma_{J1}$ is the parameter that we think is most influenced by environmental stochasticity.

@@

After parameterizing our model with CMR-estimated adult frog survival probabilities and other known vital rates (Supplementary Table 1), we performed four analyses.
First, we compute the long-run growth rate $\lambda$ for each of our 12 translocated populations to determine if the populations were deterministically predicted to grow or decline in the long-run.
Second, we compute the elasticity and sensitivity of $\lambda$ to nine model parameters to quantify how much changes in these parameters affected the long-run growth rate (Supplemental Fig. 4).
This also helped us determine where in the model environmental variation in juvenile survival and recruitment would have the largest effects on population dynamics.
Third, we included demographic stochasticity and environmental stochasticity in $\sigma_{J_1}$ in our model and simulated the 50-year viability (i.e., 1 - extinction probability) of populations given an introduction of 40 adult individuals into an unoccupied habitat.
Finally, we fit our model to our longest translocation trajectory to confirm that our model could reasonably reproduce the observed recovery trajectories of *R. sierrae* following reintroductions.
Additional details are provided in **Supporting Information - Population viability modeling - Model analysis and simulation**.

@@

![**Estimated growth rates and extinction probabilities of reintroduced populations**. **A** Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and year-1 juvenile survival and recruitment probability $\sigma_{J_1}$, given the parameterized, deterministic model. Colored points show the predicted $\lambda$ values for the twelve translocated populations when year-1 juvenile survival probability is $\sigma_{J_1} = 0.09$ (indicated by the dashed line). The red line shows where $\lambda = 1$. Note that the point for 70413 is mostly hidden behind other points. **B** Predicted 50-year extinction probabilities of the 12 reintroduced populations, given demographic stochasticity, environmental variability in $\sigma_{J_1}$, and different mean values of $\sigma_{J_1}$. There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are partially or completely hidden beneath the line for 70505. **C** 100 simulated trajectories (black lines) from the population viability model that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_manuscript.jpg){#fig-viability width="95%"}

>>>>

Given observed yearly adult survival probabilities of translocated frogs (from site-specific mrmr CMR models; provided in legend of [@fig-viability]B) and a yearly survival probability of the year-1 juvenile class ($\sigma_{J_1}$) greater than 0.10, at least six of twelve translocated populations are likely to experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd ([@fig-viability]A; median predicted $\lambda$ ranges from 1.07-1.28 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
As year-1 juvenile survival probability increased above 0.2, the deterministic long-run growth rate of seven of twelve populations was greater than 1 ([@fig-viability]A).

Even when incorporating (i) demographic stochasticity and (ii) environmental stochasticity in year-1 juvenile survival (the transition that we expect to be the most subject to environmental variability in the presence of Bd), populations with high adult survival are likely to persist over a 50 year time horizon.
Our model predicted that, following a single introduction of 40 adult individuals into a population, the six populations with the highest adult survival probabilities ($\sigma_{A_R} > 0.5$) had 50-year extinction probabilities of less than 0.5 when the average year-1 juvenile survival was greater than 0.10 ([@fig-viability]B).
This indicates strong potential for long-term persistence in the presence of Bd and environmental variability in juvenile survival.
In contrast, for the six populations where yearly adult survival probability $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be \> 50% regardless of the value of mean year-1 juvenile survival between 0 and 0.25.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history ([@fig-viability]C; population 70550, surveyed for 16 years).

@@

The recent emergence of Bd worldwide has contributed to the decline of hundreds of amphibian species, some of which are now extinct in the wild [@scheele2019].
This extraordinary impact on global amphibian biodiversity is compounded by the lack of any effective and broadly applicable strategies to reverse these impacts [@garner2016; @knapp2022].
Importantly, in addition to the natural recovery documented for *R. sierrae* [@knapp2016], other amphibian species are also showing evidence of post-epizootic recovery in the presence of Bd [@scheele2017; @voyles2018] and suggest the possibility of also using animals from these recovering populations to reestablish extirpated populations.
As with *R. sierrae*, the feasibility and long-term success of such efforts will depend on the availability of robust donor populations containing individuals that have sufficiently high resistance to allow frog survival and population growth in the presence of Bd.
Despite the promising example provided by our study of successful reestablishment of *R. sierrae* despite ongoing Bd infection, the challenge of recovering hundreds of Bd-impacted amphibian species globally is a daunting prospect.
An adaptive management approach such as the one we utilized, in which the outcomes of recovery actions are rigorously assessed and this information is used to reduce uncertainties and inform future actions, will be essential in maximizing the effectiveness of this global recovery effort [@canessa2019].
In addition, developing an improved understanding of the factors facilitating or limiting the development of frog resistance against Bd infection and the effectiveness of selective breeding programs and other genetic interventions to enhance frog resistance [@kosch2022] may be particularly important.
Although we now have a proven strategy to reestablish extirpated *R. sierrae* populations, recovery across their large historical range will require substantial resources over many decades.
The results of this study provide a hopeful starting point for that endeavor and other future efforts worldwide.

@@

To determine the implications of observed 1-year adult survival on the long-term viability of populations established via translocation, we developed a population model for *R. sierrae*.
Our central question was: How does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations?
We developed a model that tracked seven state variables of a frog population: abundance of translocated adults ($A_T$), abundance of adults naturally recruited into the population ($A_R$), abundance of first-year tadpoles ($L_1$), abundance of second-year tadpoles ($L_2$), abundance of third-year tadpoles ($L_3$), abundance of first-year juveniles ($J_1$), and abundance of second-year juveniles ($J_2$).
We divided adults into two classes $A_T$ and $A_R$ because there is evidence that the survival probability of translocated adults and naturally recruited adults differs (Supplementary Fig. 3).
The modeled population included both male and female frogs.

@@

$$
\begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\ 
A_T
\end{bmatrix}(t + 1) = 
\begin{bmatrix}
  0 & 0 & 0 & 0 & 0 & \sigma_{A_R} p_F \rho F & \sigma_{A_T} p_F \rho F \\
  \sigma_{L1} p_{L1} & 0 & 0 & 0 & 0 & 0 & 0 \\
  0 & \sigma_{L2} p_{L2} & 0 & 0 & 0 & 0 & 0\\
  \sigma_{L1} (1 - p_{L1}) & \sigma_{L2} (1 - p_{L2}) & \sigma_{L3} & 0 & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} p_{J1} & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} (1 - p_{J1}) & \sigma_{J2}  & \sigma_{A_R} & 0 \\
  0 & 0 & 0 & 0 & 0 & 0 & \sigma_{A_T} 
\end{bmatrix} \begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\
A_T \\
\end{bmatrix}(t) 
$$

The parameters in this model are yearly survival probability $\sigma_{\cdot}$ (the subscript "$\cdot$" indicates a particular state variable), probability that a female frog reproduces in a given year \$p_F\$, the percentage of adults that are female $\rho$, number of eggs produced by a female frog in a year that successfully hatch $F$, probability of a first-year tadpole remaining as a tadpole $p_{L1}$, probability of a second-year tadpole remaining as a tadpole $p_{L2}$, and probability of a first-year juvenile remaining as a juvenile $p_{J1}$.
First-year juvenile survival $\sigma_{J1}$ is the parameter that we think is most influenced by environmental stochasticity.

@@

After parameterizing our model with CMR-estimated adult frog survival probabilities and other known vital rates (Supplementary Table 1), we performed four analyses.
First, we compute the long-run growth rate $\lambda$ for each of our 12 translocated populations to determine if the populations were deterministically predicted to grow or decline in the long-run.
Second, we compute the elasticity and sensitivity of $\lambda$ to nine model parameters to quantify how much changes in these parameters affected the long-run growth rate (Supplemental Fig. 4).
This also helped us determine where in the model environmental variation in juvenile survival would have the largest effects on population dynamics.
Third, we included demographic stochasticity and environmental stochasticity in $\sigma_{J_1}$ in our model and simulated the 50-year viability (i.e., 1 - extinction probability) of populations given an introduction of 40 adult individuals into an unoccupied habitat.
Finally, we fit our model to our longest translocation trajectory to confirm that our model could reasonably reproduce the observed recovery trajectories of *R. sierrae* following reintroductions.
Additional details are provided in **Supporting Information - Population viability modeling - Model analysis and simulation**.

@@

![**Estimated growth rates and extinction probabilities of reintroduced populations**. **A** Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and year-1 juvenile survival probability $\sigma_{J_1}$, given the parameterized, deterministic model. Colored points show the predicted $\lambda$ values for the twelve translocated populations when year-1 juvenile survival probability is $\sigma_{J_1} = 0.10$ (indicated by the dashed line). The red line shows where $\lambda = 1$. Note that the point for 70413 is mostly hidden behind other points. **B** Predicted 50-year extinction probabilities of the 12 reintroduced populations, given demographic stochasticity, environmental variability in $\sigma_{J_1}$, and different mean values of $\sigma_{J_1}$. There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are partially or completely hidden beneath the line for 70505. **C** 100 simulated trajectories (black lines) from the population viability model that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_manuscript.jpg){#fig-viability width="95%"}
